---
title: 'Advanced Data Analysis with R: Assignment'
author: "Weiyi Deng"
date: "October 21, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

(a) 160 respondents suffered from heart disease; 302 did not.

```{r prep, echo=FALSE}
library("euR")
data("heart")
heart_data <- as.data.frame(heart)
heart_data$ycat <- as.factor(heart_data$y)
levels(heart_data$ycat) <- c("no", "yes")
```

```{r prep11}
summary(heart_data$ycat)
```

(b) Summary of heart data:

```{r prep2, echo=FALSE}
summary(heart_data)
```

People with head disease tend to have higher systolic blood pressure, more tobacco consumption, have higher ldl, more adiposity and obesity, more type A behavior and more alcohol consumption.

```{r scatterplot, echo=FALSE}
plot(heart_data[,-c(1,11)], main="heart", col = heart_data$ycat)
```

(c) Correlation between obesity and adiposity is 0.717, which is larger than 0.7. Therefore there is concern for multicollinearity.

```{r corr, echo=FALSE}
cor(heart_data[,-c(1,11)])
```

(d) Difference in deviance between the model and the null model is 123.97. Difference in degrees of freedom is 9. P-value for the change in deviance is 2.2e-16

```{r mod1, echo=FALSE}
mod1 <- glm(y ~ .-ycat, family = binomial(link = "logit"), 
            data = heart_data)
summary(mod1)
```

```{r mod2, echo=FALSE}
mod0 <- glm(y ~ 1, family = binomial(link = "logit"), 
            data = heart_data)
anova(mod0, mod1, test="Chisq")
```

(e) Tobacco consumption, ldl, family history, type A behaviour, and age are positively related to the probabily of having heart disease.

```{r hist1, echo=FALSE}
pred <- predict(mod1, newdata = heart_data, type = "response")
yes_heart <- pred[heart_data$ycat=="yes"]
no_heart <- pred[heart_data$ycat=="no"]
hist(yes_heart, main = "distr of fitted response values of people with heart disease", breaks = 20)
hist(no_heart, main = "distr of fitted response values of people without heart disease", breaks = 20)
```

The model tend to predict people without heart disease to have a lower probability of having the disease, 
and people with heart disease to have a higher probability of having the disease. The true positive rate 
is not very high, but overall the model has reasonable performance in predicting heart disease.

## Question 2

(a) There are no missing values; data are entered as 0s and 1s.

```{r prep3, echo=FALSE}
data("Scotch")
summary(Scotch)
```

(b) Chivas.Regal is the most popular whiskey; the.Singleton the least popular.
```{r brand, echo=FALSE}
consume_by_brand <- apply(Scotch,2,sum)
consume_by_brand
barplot(consume_by_brand, las = 2, main = "number of consumptions")
consume_by_customer <- apply(Scotch,1,sum)
```

(c) The mean of whiskey consumption of cumstomers is 2.292606. The minimum consumption is 1 and maximum consumption is 20.

```{r stats}
mean(consume_by_customer)
max(consume_by_customer)
min(consume_by_customer)
```

```{r customer1, echo=FALSE}
# consume_by_customer <- apply(Scotch,1,sum)
barplot(table(consume_by_customer), xlab = "number of whiskeys consumptions", ylab = "number of customers",
        main = "histogram of consumption by customers")
```

(e) "Johnnie.Walker.Red.Label", "Glenlivet" and "Passport" are chosen as medoids. The cluster sizes are 6, 4 and 11 respectively. 
```{r dist, echo=FALSE}
Jaccard_distance <- dist(t(Scotch), method = "binary")              # take transpose of Scotch
library("cluster")
pm <- pam(Jaccard_distance, k = 3)
```

```{r dist2}
pm$medoids
table(pm$cluster)
pm$clustering
```

(f) Prefer the clustering with 3 clusters; the cluster sizes are more balanced
```{r dist3, echo=FALSE}
pm2 <- pam(Jaccard_distance, k = 4)
```

```{r dist4}
pm2$medoids
table(pm2$cluster)
pm2$clustering
```

## Question 3

Test with Prestige data:

Compare p_value by permutation test with the cor.test() result:

```{r func1, echo=FALSE}
# write a function to compute p value of correlation between simulated variables using permutation
permTest <- function(x, y, nperm_set = 1000){
  nperm <- nperm_set
  if (!is.numeric(x)||!is.numeric(y))                      # check if inputs are numeric
    stop("x and y must be numeric")
  if (length(x)!=length(y))                                # check if inputs have the same sizes
    stop("x and y must have the same length")
  obs_cor <- cor(x,y)
  cor_perm <- rep(NA,nperm)
  set.seed(401716)
  for (k in 1:nperm){
    yPerm <- sample(y,length(y),replace=FALSE)
    cor_perm[k] <- cor(yPerm,x)
  }
  if (obs_cor >0){                                        # get number of permuted correlations more extreme than observed cor
    num_extreme <- sum(cor_perm>obs_cor)
  } else {
    num_extreme <- sum(cor_perm<obs_cor)
  }
  perm_p_val <- 2*num_extreme/nperm                       # calculate the two-sided permutation p-value
  # print(perm_p_val)                                       # print out permutation p-value
  return(list("computed correlation" = obs_cor, "permutation p-value" = perm_p_val,            
              "vector of nperm correlations" = cor_perm))                         
}

# test with Prestige data
data("Prestige")
output <- permTest(Prestige$women,Prestige$prestige)                                   
cor_val <- output[[1]]
vec_nperm_cor <- output[[3]]

perm_p_value <- output[[2]]
```

```{r show1}
perm_p_value
cor.test(Prestige$women,Prestige$prestige)                            
```

```{r plot1, echo=FALSE}
plot(density(vec_nperm_cor), 
     main = "density plot for testing correlation", xlab = "correlation", ylab = "pdf")
abline(v = cor_val, col = "red")
```

Test with cityweather data:

Compare p_value by permutation test with the cor.test() result:

```{r func2, echo=FALSE}
data("cityweather")
output <- permTest(cityweather$AvgCold,cityweather$RainDays,nperm_set = 10000)                                   
cor_val <- output[[1]]
vec_nperm_cor <- output[[3]]

perm_p_value <- output[[2]]
```

```{r show2}
perm_p_value
cor.test(cityweather$AvgCold,cityweather$RainDays)                            
```

```{r plot2, echo=FALSE}
plot(density(vec_nperm_cor), 
     main = "density plot for testing correlation", xlab = "correlation", ylab = "pdf")
abline(v = cor_val, col = "red")
```